---
title: 'Class 6: R Functions'
author: 'Jaime Bruce (PID: A16568558)'
date: "2/3/2022"
output:
  pdf_document: default
  html_document: default
---

Today we will explore R functions.

We will start with calculating a grade for these example students.

```{r}
# Example input vectors to start with
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
student3 <- c(90, NA, NA, NA, NA, NA, NA, NA)
```

We could use the mean() function to calculate an average

```{r}
mean(student3, na.rm = TRUE)
```

How does the is.na() function work? Let's try it out on student2.

```{r}
student2
is.na(student2)
```
We can use this result to get at our NA values (i.e. the TRUE positions)
```{r}
student2[is.na(student2)]
```

```{r}
x <- 1:5
x
x[x > 2] <- 100
x
```

Put these two things together - mean() is.na() function.

```{r}
student2[ is.na(student2) ] <- 0
mean(student2)
```

```{r}
student3[ is.na(student3) ] <- 0
mean(student3)
```

Now we want to find the lowest score...

```{r}
student1
which.min(student1)
```
The which.min() function tells us where the smallest value is. 

```{r}
x <- student1[ -which.min(student1) ]
mean(x)
```

```{r}
x <- student2
x[ is.na(x) ] <- 0
x <- x[ -which.min(x) ]
mean(x)
```



We are close to our working code snipet that will be th body of our first function

```{r}
# First set NA values to zero
x[ is.na(x) ] <- 0
# Remove lowest score and calculate average
mean( x[-which.min(x) ] )
```

Now we can turn this into our first function. We will call this function 'grade()'

All R functions have 3 things
- a name (grade)
- input arguments (student scores)
- body (does the work!)

```{r}
grade <- function(x) {
  # First set NA values to zero
  x[ is.na(x) ] <- 0
  # Remove lowest score and calculate average
  mean( x[-which.min(x) ] )
}
```

Le'ts try it out
```{r}
grade(student3)
```


>Q1. Write a function grade() to determine an overall grade from a vector of student homework assignment scores dropping the lowest single score. If a student misses a homework (i.e. has an NA value) this can be used as a score to be potentially dropped. Your final function should be adquately explained with code comments and be able to work on an example class gradebook such as this one in CSV format: “https://tinyurl.com/gradeinput” [3pts]

```{r}
gradebook <- read.csv("https://tinyurl.com/gradeinput", row.names=1)
head(gradebook)
```

We can use the 'apply()' function to grade the whole class.

```{r}
scores <- apply(gradebook, 1, grade)
```


>Q2. Using your grade() function and the supplied gradebook, Who is the top scoring student overall in the gradebook? [3pts]

```{r}
which.max(scores)
```

>Q3. From your analysis of the gradebook, which homework was toughest on students (i.e. obtained the lowest scores overall? [2pts]

```{r}
hw.mean <- apply(gradebook, 2, mean, na.rm=TRUE)
which.min( hw.mean )
```
```{r}
boxplot(gradebook)
```



